\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{mathtools}

\title{LP Investigation Notes}
\author{}
\date{}

\begin{document}

\maketitle

\section{What the current obstruction really is}

The issue described in \texttt{2\_current\_status.tex} is not just that the witness
sets use too few shifts. The deeper problem is that the present LP is built from a
one-parameter family of events
\[
t \in dA-x \iff \frac{t+x}{d} \in A.
\]
So each base set should be viewed as an affine form in one random parameter:
\[
L_{d,x}(t) := \frac{t+x}{d}.
\]

Then an atom is recording the joint membership pattern of the values
$1_A(L_{d,x}(t))$.

In this language, a forbidden triple appears exactly when
\[
L_i(t)+L_j(t)-mL_k(t)\equiv 0
\]
as an affine identity. Equivalently,
\[
\frac{1}{d_i}+\frac{1}{d_j}=\frac{m}{d_k},
\qquad
\frac{x_i}{d_i}+\frac{x_j}{d_j}=\frac{mx_k}{d_k}.
\]
This is the clean generalization of the $m=4$ condition $x_i+x_j=2y_k$.

\section{What changes in the affine-form prototype}

The file \texttt{affine\_form\_lp.py} implements this formulation directly.
Compared with \texttt{search\_witness\_dil.py}, it does two things differently:

\begin{itemize}
\item It treats $dA-x$ as the affine form $L_{d,x}(t)=(t+x)/d$, so the forbidden
triples are derived from the exact identity $L_i+L_j-mL_k=0$.
\item It groups subset constraints by full affine reparameterization of the
parameter $t$, rather than by ad hoc translation plus a few extra dilation
equalities.
\end{itemize}

So this is a mathematically cleaner LP, and it works for any $m$.

\section{What small experiments say}

Small tests with the new builder gave:

\begin{itemize}
\item $m=3$, layers $[(1,[0,1]), (2,[0,1])]$: bound $0.5$.
\item $m=3$, layers $[(1,[0,1,2]), (2,[0,1,2])]$: bound $0.5$.
\item $m=4$, layers $[(1,[0,1]), (2,[0,1])]$: bound $0.5$.
\item $m=4$, layers $[(1,[0,1,2]), (2,[0,1,2])]$: bound $0.5$.
\item Small brute-force searches over dilations like $(1,2,3)$ and $(1,2,4)$ with
2--3 shifts per layer also stayed at $0.5$.
\end{itemize}

So the conclusion is that the present failure is not caused by the particular
translation or dilation equalities in \texttt{search\_witness\_dil.py}. Even after
replacing them by the exact one-parameter affine symmetry, the LP still has too
much room.

\section{Promising LP directions from here}

\subsection*{A. Two-parameter linear-form LP}

This is the most promising extension.

Instead of sampling one random parameter $t$, sample two parameters $(u,v)$ and
look at linear forms
\[
L_r(u,v)=a_ru+b_rv+c_r.
\]
Then a forbidden triple is encoded whenever
\[
L_i+L_j-mL_k\equiv 0.
\]

This removes the main obstruction from \texttt{2\_current\_status.tex}: cancellation
no longer has to come from the single common variable $t$; it can happen in the
coefficient vectors $(a_r,b_r)$.

For example, for general $m$ one can take
\[
L_1(u,v)=u,\qquad
L_2(u,v)=v,\qquad
L_3(u,v)=\frac{u+v}{m}.
\]
Then $L_1+L_2-mL_3=0$ identically.

\subsection*{B. Extension-hierarchy LP}

Keep the one-parameter affine-form model, but do not optimize on a single witness
family in isolation. Introduce variables for a larger witness family $F'$ and
require that its atom distribution marginalizes to the atom distribution on $F$.
This gives a hierarchy of local consistency constraints.

\subsection*{C. LP plus additive-combinatorial cuts}

If the pure atom LP keeps admitting fake $1/2$ pseudodistributions, add
inequalities coming from external theorems rather than symmetry alone. Natural
candidates include:

\begin{itemize}
\item Pollard-type popular sum inequalities.
\item Cauchy-Davenport lower bounds on sumsets of dense slices.
\item Robust $3k-4$ style inequalities on structured pieces.
\end{itemize}

\section{Recommended next step}

If the goal is to find a formulation that has a realistic chance to beat $1/2$,
the next thing to implement is:

\begin{enumerate}
\item A two-parameter linear-form LP for $1_A(L_r(u,v))$.
\item Optionally, a small extension hierarchy on top of that.
\end{enumerate}

If the goal is only to continue searching within the current family, use
\texttt{affine\_form\_lp.py}, not \texttt{search\_witness\_dil.py}, because it is
the cleaner formulation and already supports arbitrary $m$.

\end{document}
