\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}

\title{Experiments: Two-Parameter LP Bounds for $x+y=mz$}
\author{}
\date{}

\begin{document}

\maketitle

\section{Purpose}

This note records reproducible numerical upper bounds produced by the
two-parameter linear program in \texttt{two\_parameter\_lp.py} (see also
\texttt{latex/4\_two\_parameter\_model.tex}). These are \emph{LP relaxation}
bounds: they constrain any joint law on $\mathbb{F}_p^2$ that is consistent with
affine symmetry and with the empty-atom constraints coming from identities
$L_i+L_j-mL_k\equiv 0$. They are not proved theorems about $m$-sum-free sets
unless one supplies an additional argument that realizes the same LP from actual
sets $A\subset\mathbb{F}_p$.

\section{Witness families}

All experiments use affine forms $L_r(u,v)=a_r u+b_r v+c_r$ with rational
coefficients (implemented as \texttt{Fraction}s) and integer shifts in the
constant term. The helper \texttt{pentad\_family(m)} in
\texttt{two\_parameter\_lp.py} returns the five directions
\[
L_1=u,\quad L_2=v,\quad L_3=\frac{u+v}{m},\quad
L_4=\frac{2u-v}{m},\quad L_5=\frac{u-2v}{m},
\]
each with a single shift at $0$.

The classical ``four-direction'' family (coordinates $u$, $v$, $(u+v)/m$,
$(u-v)/m$) yields the relaxation bound $4/7$ in the initial runs recorded
earlier. Adding the two extra rational directions $(2u-v)/m$ and $(u-2v)/m$
reduces the LP value when $m=3$.

\section{Best value found for $m=3$}

Exhaustive search over all $5$-form families obtained by fixing $u$ and $v$ and
choosing any three further directions from a small integer pool (see
\texttt{experiments\_two\_parameter.py}) gives a largest observed optimum of
\[
\delta(A)\le \frac{8}{15}\approx 0.5333
\]
for the pentad above. In the same search, the four-direction
$u,v,(u+v)/3,(u-v)/3$ family gives $4/7\approx 0.5714$.

An exhaustive sweep over $6$-form families from the same pool (two coordinates
$u,v$ plus four further directions, single shift each) did not improve on $8/15$
in floating-point SciPy solves.

\section{Dependence on $m$}

For the \emph{same} pentad family, the reported LP optimum returns to $4/7$ for
sampled odd $m\ge 5$ (e.g.\ $m\in\{4,5,7\}$ in double-precision checks). Thus the
improvement to $8/15$ is not uniform in $m$; it reflects additional linear
relations among the five directions when $m=3$.

\section{Code and solver caveats}

\begin{itemize}
\item \texttt{evaluate\_two\_parameter\_bound} and \texttt{evaluate\_affine\_bound}
return \texttt{None} if \texttt{scipy.optimize.linprog} does not report success,
instead of silently returning $1/2$.
\item Very large witness families make the LP expensive; start with few forms and
increase slowly.
\end{itemize}

\section{Superseded by the three-parameter LP}

The three-parameter LP in \texttt{three\_parameter\_lp.py} improves on every
bound of this note, uniformly in $m\ge 3$. Its six-form, nine-form and
twelve-form families give respectively $16/33\approx 0.4848$,
$48/103\approx 0.4660$ and approximately $0.4398$ for $m=3$. See
\texttt{latex/6\_three\_parameter\_model.tex} and
\texttt{latex/7\_three\_parameter\_results.tex}.

\section{Ideas for stronger relaxations}

\begin{itemize}
\item \textbf{More parameters.} Three independent parameters $(u,v,w)$ and
linear forms $a u+b v+c w+d$ enlarge the symmetry group and the space of
forbidden triples; the LP size grows as $2^n$.
\item \textbf{Consistency across families.} Couple several witness families by
requiring that marginals of a single global law agree on overlaps (hierarchy /
Sherali--Adams style).
\item \textbf{Non-symmetric cuts.} Import inequalities from additive combinatorics
(Cauchy--Davenport, popular sums) that are valid for \emph{true} set laws but not
for arbitrary LP pseudodistributions.
\item \textbf{Modular constraints.} For fixed small $p$, brute-force $A$ and
compare LP value to $\max_A |A|/p$ to see the relaxation gap on that $p$.
\end{itemize}

\end{document}
