\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{booktabs}

\title{Numerical results of the three-parameter LP for $x+y=mz$}
\author{}
\date{}

\newtheorem{remark}{Remark}

\begin{document}

\maketitle

\section{Summary table}

The following upper bounds on the LP optimum (hence on the density of any
set $A\subset\mathbb{F}_p$ avoiding $x+y=mz$) were computed with
\texttt{three\_parameter\_lp.py}, using SciPy's HiGHS solver. See
\texttt{experiments\_three\_parameter.py} for an exact reproduction script.

\medskip

\noindent\begin{tabular}{l c c c l}
\toprule
Family                                           & $n$ & $m$           & Bound                & Reproducer                         \\
\midrule
One-param pentad \cite{LMPV}-style               & $5$ & $2$           & $2/7$                & \texttt{python3 run\_proof.py}      \\
Two-param four directions                        & $4$ & $3$           & $4/7$                & \texttt{experiments\_two\_parameter.py} \\
Two-param \texttt{pentad\_family}                & $5$ & $3$           & $8/15$               & \texttt{experiments\_two\_parameter.py} \\
Two-param \texttt{pentad\_family}                & $5$ & $\{4,5,7\}$   & $4/7$                & \texttt{experiments\_two\_parameter.py} \\
\midrule
Three-param \texttt{basic\_pair\_family}         & $6$ & any $\ge 3$   & $\mathbf{16/33}$     & \texttt{experiments\_three\_parameter.py} \\
Three-param \texttt{signed\_pair\_family}        & $9$ & any $\ge 3$   & $\mathbf{48/103}$    & \texttt{experiments\_three\_parameter.py} \\
Three-param \texttt{symmetric\_12\_family}       & $12$& $3$           & $\mathbf{\approx 0.43983}$ & \texttt{experiments\_three\_parameter.py} \\
Three-param \texttt{symmetric\_12\_family}       & $12$& $4$           & $\approx 0.43864$    & \texttt{experiments\_three\_parameter.py} \\
Three-param \texttt{symmetric\_12\_family}       & $12$& $\ge 5$       & $\approx 0.44322$    & \texttt{experiments\_three\_parameter.py} \\
\bottomrule
\end{tabular}

\medskip

The \textbf{bold} lines are the new bounds introduced in this project.

\section{Comparisons}

\begin{itemize}
\item The one-parameter LP from \texttt{affine\_form\_lp.py} stalls at $1/2$
for every $m\ge 3$ tested (see \texttt{latex/3\_lp\_investigation\_notes.tex}).
\item The two-parameter LP stalls at $8/15\approx 0.5333$ for $m=3$ and
$4/7\approx 0.5714$ for $m\in\{4,5,7\}$, across all witness pools we have tried
(\texttt{latex/5\_experiments\_and\_bounds.tex}).
\item The three-parameter LP with only six forms already drops below $1/2$,
giving $\delta(A)\le 16/33\approx 0.4848$ uniformly in $m\ge 3$.
\item Adding the three signed differences improves the bound further to
$48/103\approx 0.4660$, still uniform in $m$.
\item Adding the three $(2a-b-c)/m$ forms gives $\approx 0.4398$ for $m=3$ and
smaller values for $m\in\{4,5,7\}$ when computed in double precision. The
dependence on $m$ for the twelve-form family is visible but small; it comes
from identities $L_i+L_j=mL_k$ that engage the new directions only for certain
$m$.
\end{itemize}

All bounds are stable under re-solving with SciPy HiGHS on the same families
and agree with the exact rationals $16/33$ and $48/103$ to machine precision.

\section{Reproducing the bounds}

To reproduce every row in the table:
\begin{verbatim}
python3 run_proof.py
python3 experiments_two_parameter.py
python3 experiments_three_parameter.py
\end{verbatim}

Typical runtimes on a single laptop core: the six- and nine-form families
finish in well under a second each; the twelve-form family takes about
3--4 minutes per $m$.

\section{Scope and next steps}

As in the previous notes these are LP relaxation bounds for affine-invariant
atom laws; turning them into theorems for $m$-sum-free sets requires
identifying the LP with set statistics. The main open directions are:

\begin{itemize}
\item[(i)] Extend \texttt{symmetric\_12\_family} to fifteen or eighteen forms
(e.g., adding $(u+v+w)/m$, $(u+v-w)/m$ and its cyclic orbit, or
\emph{integer-translation} shifts).
\item[(ii)] Replace the single witness family by a joint family (hierarchy /
Sherali--Adams style) to enforce marginal compatibility.
\item[(iii)] Introduce non-symmetric cuts from additive combinatorics
(Cauchy--Davenport, popular sums, $3k-4$ inequalities) so the LP knows more
about the integer set structure than just affine symmetry.
\end{itemize}

\end{document}
